An anthropological account of the Vim text editor:\\ features and tweaks after 10 years of usage

### 2017-12-18 article/article.tex
added version text file

Because browsing the interface in Vim is fast,
and it favors copy and paste of text,
it is often handy to keep (e.g. a tab with) some terminals:
one with an IPython shell, another two for compiling \LaTeX
and opening PDF files (e.g. with \tttt{\$ evince <filename.pdf>}).
It is very comfortable to copy and browse the terminal lines
in the Terminal-Normal mode.
I found the mappings on~\cite{vimrc} very helpful for directing
editor focus to splits (\ttt{C-hjkl}) and tabs (\tttt{gr}, \tttt{gt}),
which I make available across Terminal-Job and Normal modes.
But I've been thinking on using \ttt{C-} commands also
for tabs (not only for splits), both to avoid mixing typing sequences
with commands, and to allow the use of the commands in all Normal,
Insert and Terminal modes.

\subsection{Netrw}\label{netrw}
The standard interface of Vim for browsing file trees is Netrw.
It starts when you open a directory, such as with \tttt{:e .<CR>}.
It has solid support for editing remote files (such as over ssh or
ftp) and handy e.g. mappings to open the files as splits and tabs (specially \tttt{pot}).
Most useful commands in netrw include:
\begin{itemize}
  \item \tttt{d} and \tttt{\%} for creating directories and files.
    \ttt{Delete} removes both.
  \item \tttt{mf} and \tttt{mb} for marking files and bookmarking directories.
  \item \tttt{gb} and \tttt{uU} are used to load directories
    while marked files might be copied, moved, edited, grep-ed, tagged and migrated
    to and from the argslist as in \tttt{:h netrw-mf}.
\end{itemize}
Last notes:
there is no insert mode in netrw interface;
the commands in \tttt{:h netrw-explore} are
convenient for opening the directory of the
file being edited;
further information in \tttt{:h netrw}.

\subsection{Standard configuration files and directories and my .vim/vimrc}
You can check the scripts Vim loads by using the debug script mentioned
in Section~\ref{state}.
By default, \tttt{$\sim$/.vimrc} and \tttt{$\sim$/.vim/vimrc} files are run by Vim at the beginning of the startup.
One might edit the vimrc file with \tttt{:e \$MYVIMRC}
and reload it with \tttt{:source \$MYVIMRC}, but
I use the mappings in~\cite{vimrc} to encourage a continuous enhancement of my settings.
Any other file might be included to run at startup by
adding a line \tttt{:source <afile.vim>} in vimrc.
In fact, it is on vimrc that one usually specifies the plugins
and plugin managers they use.
Use an \tttt{after/} folder of a directory in \tttt{:se runtimepath}
or follow some patterns described on the next section to change the
scripts and sequence of them to be loaded.

\subsection{Plugins and packages}
One can see the list of standard plugins
with \tttt{:h standard-plugin-list} command.
Any \tttt{plugin/**/*.vim} file inside a directory
listed by \tttt{:se runtimepath} will be loaded
(e.g. \tttt{.vim/plugin/something/ascript.vim}).
There are various ways to automate the installation
and enhnace the management of plugins.
By default, one has the GETSCRIPT interface (see \tttt{:h getscript}),
that downloads latest scripts from sourceforge as specified in \tttt{:h getscript-data},
and the Vimball interface, which creates and loads a Vimball for a plugin
with \tttt{:[range]MkVimball <filename> path}, where range specifies lines
that hold paths to files to be included in the \tttt{<filename>.vba} Vimball.
The Vimball can be installed in a system by \tttt{:source <filename>.vba}
or loading it at Vim startup with \tttt{\$ vim <filename>.vba}.

@@

All directories Vim looks for scripts are described
in \tttt{:h vimfiles} and are basically set by
\tttt{:se runtimepath} and conventions inside
each path therein, such as to look for \tttt{vimrc} files
and \tttt{plugin/} or \tttt{pack/} directories.
Vim scripts can be loaded conditionally,
e.g. only if a function is used as in \tttt{:h autoload-functions}.
Example of such are filetype plugins (enabled by in a \tttt{ftplugin/<filetype>.vim} file,
e.g. inside a plugin directory).
There is a number of plugin managers for Vim.
Pathogen and Vundle seem to be the most popular,
one because of its minimalism, the other because of
advanced features, e.g. for searching and installing plugins with
colon commands.

\subsection{Spell and spelllang (en and pt\_br)}
One might set the spelling language with\\ \tttt{:se spelllang=en\_us}
or \tttt{:se spelllang=pt\_br}
and toggle spell checking with \tttt{:setl spell!}.
These are used so often that one might use shortcuts as in~\cite{vimrc}.
Currently, Vim will download the files for a specific language if
not found in system.

\subsection{Scripting, Functions, Vimscript and other languages (e.g.  Python)}\label{script}
In Vimscript,
the colon commands (also Ex or command-line commands) are related through spaces,
punctuation and keywords (see \texttt{:h script}).
Scripting the Vim editor can also be accomplished using other languages,
as well documented e.g. \tttt{:h python}.
Functions are defined through colon commands and are called
inside colon commands e.g. \tttt{:call MyFunction()}
or \tttt{:echo MyFunction(4)}.
Notice that functions are not commands but might be bind to them
through colon commands e.g. \tttt{nnoremap gF :call MyFunction()<CR>}.
Executing source files is very straightforward with \tttt{:so
<filename>.vim},
and one can always use the Ex mode for rapid scripting.

At the same time, the \tttt{:term} and terminal-job mode make
scripting other software more convenient, as output is
promptly navigated and copied. The shell works well for generic use,
such as opening images and PDFs through \tttt{\$ eog} and \tttt{\$ evince}.
See Section~\ref{terminal} for further directions about the terminal.

\subsection{Color}\label{visual2}
Newer versions of Vim support 24 bit true color (aka 16 million colors)
in terminal Vim (not only in GVim).
The terminal must support true color, and tests are available e.g. in
\url{https://gist.github.com/XVilka/8346728}.
Then Vim needs to be set to use true colors with
\tttt{:set termguicolors}.
If using 8 or 16 bit colors, Vim uses the color pallete from the
terminal, if using true colors each color is defined directly.
Settings for using true colors inside Byobu/Tmux involve tweakings
and are available in~\cite{vimrc}.
Good color schemes to use with true colors are Gruvbox
and Solarize.
One might source syntax files at any time to change syntax, usually
though linking tokens to syntax groups with
\tttt{:syntax keyword <group\_name> <token1> <token2> <token3> ...}
and then relating the group to another group 
\tttt{:highlight link <group\_name> <group\_name2>}
or by specifying the group characteristics directly:
\tttt{highlight <group\_name> guifg=\#ffffff}.
If you change a syntax file, reloading a file with with \tttt{:e<CR>}
updates the highlighting on the window with the corresponding filetype.
A complete syntax highlighting support typically involves at least three files:
\begin{itemize}
  \item \tttt{ftdetect/<filetype>.vim}, where the file type is
    detected with e.g.
    \tttt{:autocmd BufNewFile,BufRead *.<file\_extension> setfiletype <filetype>}
  \item \tttt{ftplugin/<filetype>.vim} with general settings for the
    file type, such as: 
    \tttt{:set tabstop=2 softtabstop=2 shiftwidth=2 textwidth=70
    expandtab autoindent}.
  \item A \tttt{syntax/<filetype>.vim} file, with bindings between
    tokens and highlighting groups; and highlighting group definitions.
\end{itemize}

This very scheme is implemented very straightforward in this
plugin~\cite{tokipona}
for highlighting text in the Toki-Pona language~\cite{tpLang}.
Syntax highlighting plugins are as filetype plugins,
but also have a \tttt{syntax/<filetype>.vim} file
relating keywords to highlighting groups. 
One might see every highlighting group, and their final visual results,
with the command
\tttt{:so \$VIMRUNTIME/syntax/hitest.vim}.
In my system, the \tttt{ftdetect/} and \tttt{ftplugin/} folders load as expected in the \tttt{plugin/}
directory, but \tttt{syntax/} files had to be moved to \tttt{$\sim$/.vim/syntax/}.

\subsection{Fonts}
The fonts are defined by your terminal or inside GVim.
\ttt{C-+} and \ttt{C--} can be used to change font size.
Some settings for fonts, such as boldface, might be set using
the syntax highlighting facilities described in the previous section.

\section{Conclusions and further work}\label{conc}
This document seems reasonable as an overall reference of the Vim editor,
at least for my usage and proficiency.
Given the folkloric milestone of using Vim for 10 years,
this article might serve as a benchmark for one to relate
it's current use and understandings.
As a pedagogical material, it seems to be unique in the emphasis
on namespaces, understood as commands, variables, state-related lists, etc,
especially in Section~\ref{namespaces},
and the reference to the standard Vim documentation
to achieve the DRY KISS style described in Section~\ref{intro}.

\noindent {\bf Potential enhancements} to this document are:
\begin{itemize}
  \item The inclusion of facilities such as reading emails and connecting over ssh.
    There is a working hack in~\cite{vimrc} for browsing over the WWW,
    but such aspect of Vim usage might receive more attention
    given that it is confortable to use the navigation and editing facilities
    and the resulting integrated environment.
  \item Updating of the information I can find about the issues discussed,
    such as about status lines in Section~\ref{visual1}.
  \item Include a discussion about Neovim.
    I have never used it, but it seems to be reaching a considerable user base
    and it might be feasable to give an account of Vim and Neovim
    after some tests and researching the official and user base documentation.
  \item Better cite documentation, plugins and Vim authors.
    I prefered to keep the references inline through \tttt{:h} commands
    and URLs, more in accordance with the style of Vim documentation,
    but bibliographical items constitute a valuable asset for academic literature,
    and some authors might find their work more pretiged if more
    thoroughly cited.
  \item An analysis of my usage, e.g. according to \url{http://www.drbunsen.org/vim-croquet/}.
    This undertake might benefit from data from many users,
    which favors the potential plugin for usage analytics described in the next bullet list.
\end{itemize}

\noindent {\bf Potential next steps} in using Vim:
  \item Measure the performance of text mining routines in Vim against those implemented in C or Python.
  \item Enhance the HTTP browsing capabilities of~\cite{vimrc}.
  \item Better integrate Python and Vimscript, especially for data visualization
    and syntax highlighting management.
    In accordance with the visualization issues described in
    Sections~\ref{visual1} and~\ref{visual2}.
\noindent Many of the next steps might entail plugins.
This is a selection of the issues that are more prominent for me:

@@

@@

      \item Slick Vim: a collection of the settings, mappings and plugins I use.

@@

        and dummy minimal plugins as simplest, file type and syntax highlighting,

@@

        but also hold stylistic merits as ascii art is often

@@

        One can both obtain an ascii representation of a binary image (e.g. JPG, PNG),
        and can directly render ascii charts from data using cues such are shape, position

@@

      \item Redirecting the commands that usually show the results in a 'more' interface,
        which cannot be searched nor copied,
        to a parsed and linked quickfix or location window.

@@

        slides and opening figures and some Vim users asked for the settings and
        commands. They are very elementary use of registers being executed over

@@

@@

        syntax highlighting scheme and dump it to a file (dumpung
        current highlighting scheme is currently not supported).
        Should also include changing the color scheme an highlighting scheme incrementally
        and selectively using the features described in Section~\ref{visual2}.
      \item Run Python excepts from file in an IPython shell. 
        Mappings should make available all movements to catch the excepts,
        execution of registers, remain on script or in the IPython shell.
        Also, terminal mode can use some improvements, such as \ttt{C-O} for
        one normal mode command.
        for optimizing the usage, as described in~\cite{http://www.drbunsen.org/vim-croquet/}. Usage analytics.
      \item Colorschemes for true color (gui).

@@

        e.g. SpellBad tokens are not highlighted on these colorschemes
        it you have a functioning \tttt{se termguicolors} (standard GUI mode
        and available in terminals since recent versions of Vim 8, see Section~\ref{visual2}). In~\cite{vimrc} are some lines that make such colorschemes after over

@@

      \item Opening the editor status listings such as in Section~\ref{}
        in a quickfix or location window where they might be searched, edited
        and used as wanted.
        These windows open in a more-like interface that is very restricted in what
        they allow (e.g. no search, no persistence).

@@

\appendix
\section{Example of usage session}
I usually begin by opening a file or directory
with \tttt{\$ vim <filename>}.
The color scheme is alternated between
blue and GruvBox with
\tttt{:colo gruvbox} and \tttt{:colo blue}.
I open an vertical split and then move
the window to a new tab using
\tttt{:vs} and \ttt{C-W\_T}.
I then search tokens related to
the enhancements I want to make or
the knowledge I want to acquire.
I go back to the previous tab with \tttt{gr}
and make a global replace with
\tttt{:\%s/<this>/<that/g}.
On adding dots to sentences,
I record in the \tttt{"q} register
the sequence \tttt{jA.jj},
using it as a macro 10 times by
\tttt{10@q}.
I move to the other tab with \tttt{gt}
and open a terminal window with \tttt{:term}
for compiling latex files.
I start another terminal with \tttt{<C-W> :term}
for opening the resulting PDFs with evince.
If any new idea comes to mind and I have time,
\tttt{\textbackslash\textbackslash s} opens 
my vimrc~\cite{vimrc} for editing and \tttt{\textbackslash s}
sources it.
If there is e.g. code or notes in other projects I am working
on, I reach them through \tttt{<Alt-arrows>} as other Vim instances
in Byobu/Tmux sessions and windows.

\section{My vimrc file and usage}
In this Appendix is my vimrc file.
Although it has comments, one should
look for the options that (s)he does
not understand, as a thorough explanation
of the file is tedious and out of the scope
of this document.

In using Vim with my vimrc file~\cite{vimrc},
I mostly toggle the status line with \texttt{\textbackslash\textbackslash B}
and the tab line with \texttt{\textbackslash\textbackslash T}.
Save and close windows with \texttt{\textbackslash w} and \texttt{\textbackslash q}.
The mappings for transitioning through splits and tabs
are also used constantly.

@@

\subsection{Normal mode}
Every letter and character in the keyboard is used.
In Normal mode: \ttt{TAB} is the same as \ttt{C-I},
\ttt{BS} and \ttt{C-H} is the same as h.
\ttt{C-J} and \ttt{C-N} is the same as j.
\ttt{C-P} is the same as k.
Space is same as l.
\ttt{C-[} and \ttt{Esc} are not used,
\tttt{<C-\textbackslash> a-z} is reserved for extensions,
\ttt{C-\_} not used.
\tttt{+} is the same as \ttt{CR} and they are both not very useful.
Del is same as x.

@@

Directions, home, end, page up and down, insert, all have mappings
in more centraly located keys.
unbounded: g,z,[]

@@

\subsection{Insert mode}
All standard char keys are used for entering chars on text.
\tttt{<C-G>(j,k)} can be achieved by \tttt{<C-O>(j,k)} which is more powerful
in moving through multiple lines.
\ttt{C-[} is the same as \ttt{ESC}.
\ttt{C-J} and \ttt{C-M} are \ttt{CR}.
\tttt{<C-\textbackslash> a-z} is reserved for extensions,
other combinations with \ttt{C-\textbackslash} are not used.

>>>>

Because browsing the interface in Vim is fast,
and it is very comfortable to copy the terminal lines
in the Terminal-Normal mode,
it is often handy to keep (e.g. a tab with) some terminals:
e.g. one with an IPython shell, another two for compiling \LaTeX\
and opening PDF files (e.g. with \tttt{\$ evince <filename.pdf>}).
There are even more convenient ways to use the terminal inside Vim.
For example, 
one might use \tttt{:term} with the \tttt{++hidden} and \tttt{++close}
options to compile a \LaTeX\ file in the background without needing to further manage the
terminal and in a non-blocking manner: \tttt{:term ++hidden pdflatex
\% \#}.
I found the mappings on~\cite{vimrc} very helpful for directing
editor focus to splits (\ttt{C-hjkl}) and tabs (\tttt{gr}, \tttt{gt}),
which I make available across Terminal-Job and Normal modes.
But I've been thinking on using \ttt{C-} commands also
for tabs (not only for splits), both to avoid mixing typing sequences
with commands, and to allow the use of the commands in all Normal,
Insert and Terminal modes.

\subsection{Netrw}\label{netrw}
The standard interface of Vim for browsing file trees is Netrw.
It starts when you open a directory, such as with \tttt{:e .<CR>}.
It has solid support for browsing remote file trees (such as over ssh or
ftp) and handy e.g. mappings to open the files as splits and tabs
(specially \tttt{pot}).
Most useful commands in netrw include:
\begin{itemize}
  \item \tttt{d} and \tttt{\%} for creating directories and files.
    \ttt{Delete} removes both.
  \item \tttt{mf} and \tttt{mb} for marking files and bookmarking directories.
  \item \tttt{gb} and \tttt{uU} are used to load directories
    while marked files might be copied, moved, edited, grep-ed, tagged and migrated
    to and from the argslist as in \tttt{:h netrw-mf}.
\end{itemize}

There is no insert mode in netrw interface;
the commands in \tttt{:h netrw-explore} are
convenient for opening the directory of the
file being edited;
further information is in \tttt{:h netrw}.

\subsection{Standard configuration files and directories and my .vim/vimrc}
You can check the scripts Vim loads by using the debug script mentioned
in Section~\ref{state}.
By default, \tttt{$\sim$/.vimrc} and \tttt{$\sim$/.vim/vimrc} files are run by Vim at the beginning of the startup.
One might edit the vimrc file with \tttt{:e \$MYVIMRC}
and reload it with \tttt{:source \$MYVIMRC}.
Mappings in~\cite{vimrc} include such commands
to encourage a continuous enhancement of Vim settings
(they have helped me to improve my settings without unnecessary 
hassle and fast).
Any other file might be included to run at startup by
adding a line \tttt{:source <afile.vim>} in vimrc.
In fact, it is on vimrc that one usually specifies the plugins
and plugin managers they use.
Use an \tttt{after/} folder of a directory in \tttt{:se runtimepath}
or follow some patterns described on the next section to change the
scripts and sequence of them to be loaded.
The vimrc from other users are most useful for one to comprehend
and pick convenient practices and settings.
In fact, a vimrc file is most often a collage of excerpts of vimrc
files from other users.

\subsection{Plugins and packages}
One can see the list of the standard plugins
with \tttt{:h standard-plugin-list} command.
Any \tttt{plugin/**/*.vim} file inside a directory
listed in \tttt{:se runtimepath} will be loaded
(e.g. \tttt{.vim/plugin/something/ascript.vim}).
There are various ways to automate the installation
and enhance the management of plugins.
By default, one has the GETSCRIPT interface (see \tttt{:h getscript}),
that downloads latest scripts from Sourceforge as specified in \tttt{:h getscript-data},
and the Vimball interface, which creates and loads a Vimball for a
plugin.
Such a Vimball may be created with \tttt{:[range]MkVimball <filename> path}, where range specifies lines
that hold paths to files to be included in the \tttt{<filename>.vba} Vimball.
The Vimball can be installed in a system by \tttt{:source <filename>.vba}
or loading it at Vim startup with \tttt{\$ vim <filename>.vba}.

@@

All directories Vim looks for scripts are described
in \tttt{:h vimfiles} and are basically set by
\tttt{:se runtimepath} and conventions inside
each directory therein, such as to look for \tttt{vimrc} files
and \tttt{plugin/} or \tttt{pack/} directories.
Vim scripts can be loaded conditionally,
e.g. only if a function is used as in \tttt{:h autoload-functions}.
Example of such are filetype plugins (enabled by a \tttt{ftplugin/<filetype>.vim} file,
e.g. inside a plugin directory).
There is a number of plugin managers for Vim.
Pathogen and Vundle seem to be the most popular,
one because of its minimalism, the other because of
advanced features, e.g. for searching and installing plugins with
colon commands.

\subsection{Spell and spelllang (en and pt\_br)}
One might set the spelling language with\\ \tttt{:se spelllang=en\_us}
or \tttt{:se spelllang=pt\_br}
and toggle spell checking with \tttt{:setl spell!}.
Depending on the activities being performed,
these commands are used so often that one might use mappings as in~\cite{vimrc}.
Currently, Vim will download the files for a specific language if
not found in system.

\subsection{Scripting, Functions, Vimscript and other languages (e.g.  Python)}\label{script}
In Vimscript,
the colon commands (also Ex commands or command-line commands) are related through spaces,
punctuation and keywords (see \texttt{:h script}).
Scripting the Vim editor can also be accomplished using other languages,
as well documented e.g. in \tttt{:h python}.
Functions are defined through colon commands and are called
inside colon commands e.g. \tttt{:call MyFunction()}
or \tttt{:echo MyFunction(4)}.
Notice that functions are not commands but might be bind to them
through colon commands e.g. \tttt{nnoremap gF :call MyFunction()<CR>}.
Executing source files is very straightforward with
\tttt{:so <filename>.vim},
and one can always use the Ex mode for rapid scripting.
At the same time, the \tttt{:term} and terminal-job mode make
scripting other software more convenient, as output is
promptly navigated and copied (as discussed in Section~\ref{terminal}).

\subsection{Color}\label{visual2}
Newer versions of Vim support 24-bit true color (aka 16 million colors)
in terminal Vim (in gVim true color received support earlier).
The terminal must support true color, and tests are available e.g. in
\url{https://gist.github.com/XVilka/8346728}.
Then Vim needs to be set to use true colors with
\tttt{:set termguicolors}.
If using 8 or 16 bit colors, Vim uses the color palette from the
terminal, if using true color each color is defined directly.
Settings for using true color inside Byobu/Tmux involve tweaking
and are available in~\cite{vimrc}.
Good color schemes to use with true color are Gruvbox
and Solarize.
One might source syntax files at any time to change syntax, usually
though linking tokens to syntax groups as in
\tttt{:syntax keyword <group\_name> <token1> <token2> <token3> ...},
or
\tttt{:syntax match <group\_name>	<pattern>},
and then relating the group to another group: 
\tttt{:highlight link <group\_name> <group\_name2>}
or to group characteristics directly:
\tttt{highlight <group\_name> guifg=\#ffffff}.
If you change a syntax file, reloading a file with with \tttt{:e<CR>}
updates the highlighting on the window with the corresponding file type.
A complete syntax highlighting support typically involves at least three files:
\begin{itemize}
  \item \tttt{ftdetect/<filetype>.vim}, where the file type is
    detected with e.g.
    \tttt{:autocmd BufNewFile,BufRead *.<file\_extension> setfiletype <filetype>}
  \item \tttt{ftplugin/<filetype>.vim} with general settings for the
    file type, such as: 
    \tttt{:set tabstop=2 softtabstop=2 shiftwidth=2 textwidth=70
    expandtab autoindent}.
  \item A \tttt{syntax/<filetype>.vim} file, with bindings between
    tokens and highlighting groups; and highlighting group definitions.
\end{itemize}

This scheme is implemented very straightforward in this
plugin~\cite{tokipona}
for highlighting text in the Toki Pona language~\cite{tpLang}.
Syntax highlighting plugins are as file type plugins,
but also have a \tttt{syntax/<filetype>.vim} file
relating keywords, matches and regions to highlighting groups. 
One might see every highlighting group, and their final visual results,
with the commands
\tttt{:so \$VIMRUNTIME/syntax/hitest.vim}
or \tttt{:so hi}.
The \tttt{ftdetect/} and \tttt{ftplugin/} folders load as expected in the \tttt{plugin/}
directory, but \tttt{syntax/} files has to be moved to
\tttt{$\sim$/.vim/syntax/}, unfortunately.

\subsection{Fonts}
The fonts are defined by your terminal or inside gVim.
\ttt{C-+} and \ttt{C--} can be used to change font size.
Some settings for fonts, such as boldface, might be set using
the syntax highlighting facilities described in the previous section.

\section{Conclusions and further work}\label{conc}
This document seems reasonable as an overall reference of the Vim editor,
at least for my usage and level of proficiency.
Given the folkloric milestone of using Vim for 10 years,
this article might serve as a benchmark for one to relate
it's current use and understandings.
As a pedagogical material, it seems to be unique in the emphasis
on namespaces, understood as commands, variables, state-related lists, etc,
especially in Section~\ref{namespaces},
and the reference to the standard Vim documentation
to achieve the DRY KISS style described in Section~\ref{intro}.

\noindent {\bf Potential enhancements} to this document include:
\begin{itemize}
  \item The discussion of facilities such as reading emails and connecting over ssh.
    There is a working hack in~\cite{vimrc} for browsing over the WWW,
    but such aspect of Vim usage might receive more attention
    given that it is comfortable to navigate and edit in Vim
    and the resulting integrated environment.
  \item Updating of the information I can find about the issues discussed,
    such as about status lines in Section~\ref{visual1}.
  \item Include a discussion about Neovim.
    I have never used it, but it seems to be reaching a considerable user base
    and it might be feasible to give an account of Vim and Neovim
    after some tests and researching the official and unofficial documentation.
  \item Better cite documentation, plugins and Vim authors.
    I preferred to keep the references inline through \tttt{:h} commands
    and URLs, more in accordance with the style of Vim documentation,
    but bibliographical items constitute a valuable asset for academic literature,
    and some authors might find their work more respected if more
    thoroughly cited.
  \item An analysis of my usage, e.g. according to \url{http://www.drbunsen.org/vim-croquet/}.
    This undertake might benefit from data from many users,
    which favors the potential plugin for usage analytics described in the next bullet list.
\end{itemize}

\noindent {\bf Potential next steps} in using Vim include:
\begin{itemize}
  \item Measure the performance of text mining routines in Vimscript against those implemented in C or Python.
  \item Enhance the HTTP browsing capabilities of~\cite{vimrc}.
  \item Better integrate Python and Vimscript, especially for data visualization
    and syntax highlighting management,
    in accordance with the visualization issues described in
    Sections~\ref{visual1} and~\ref{visual2}.
\end{itemize}

\noindent This is a selection of the issues that might entail
plugins and that are more prominent for me:

@@

        Maybe already group possibilities by criteria such as length
        of sequence and how central are the keys.

@@

      \item Slick Vim: a collection of the settings and mappings I use.

@@

        and dummy minimal plugins as simplest possible, then file type
        and then syntax highlighting,

@@

        but also hold stylistic merits as ASCII art is often

@@

        One can both obtain an ASCII representation of a binary image (e.g. JPG, PNG),
        and can directly render ASCII charts from data using cues such are shape, position

@@

      \item Redirecting the commands that usually show the results in
        a 'more' interface (e.g. status listings such as in
        Section~\ref{state}),
        which cannot be searched nor copied nor persists if one
        returns to editing a file.
        Ideally, it should parsed and linked quickfix or location
        window, and the syntax highlighting maintained.

@@

        slides and opening figures. Some Vim users asked for the settings and
        commands. They are very elementary use of registers which are executed over

@@

      \item Run Python excerpts from a file being edited in an IPython shell. 
        Mappings should make available all movements to fetch the
        excerpts, execute registers, remain on script or in the IPython shell.
        Also, current Terminal-Job mode can be improved easily with
        mappings, such as \ttt{C-O} for one normal mode command.
        for optimizing the usage, as described in~\cite{http://www.drbunsen.org/vim-croquet/}.
        Usage analytics.

@@

        syntax highlighting scheme and dump it to a file (dumping
        current highlighting scheme to a working Vimscript file
        is currently not supported by Vim!).
        Should also include changing the color scheme and highlighting scheme incrementally
        and selectively using the features described in Section~\ref{visual2}
        and in~\cite{tokipona}.
      \item Color schemes for true color.

@@

        e.g. SpellBad tokens are not highlighted on these color schemes
        if you have a functioning \tttt{se termguicolors} (standard GUI mode
        and available in terminals since recent versions of Vim 8,
        see Section~\ref{visual2}).
        In~\cite{vimrc} are some lines that make such color schemes over

@@

        Such color schemes should also make use of the discussions
        in~\cite{tokipona}.

@@

\appendix
\section{Example of usage session}
I usually begin by opening a file or directory
with \tttt{\$ vim <filename>}.
The color scheme is alternated between
blue and GruvBox with
\tttt{:colo gruvbox} and \tttt{:colo blue}.
I open a vertical split and then move
the window to a new tab using
\tttt{:vs} and \ttt{C-W\_T}.
I then search for tokens related to
the enhancements I want to make or
the knowledge I want to acquire.
I go back to the previous tab with \tttt{gr}
and make a global replace with
\tttt{:\%s/<this>/<that>/g}.
On adding dots to sentences,
I record in the \tttt{"q} register
the sequence \tttt{jA.jj},
and use it as a macro 10 times by
\tttt{10@q}.
I move to the other tab with \tttt{gt}
and open a terminal window with \tttt{:term}
for compiling latex files.
I start another terminal with \tttt{<C-W> :term}
for opening the resulting PDFs with evince.
If any new idea comes to mind and I have time,
\tttt{\textbackslash\textbackslash s} opens 
my vimrc~\cite{vimrc} for editing and \tttt{\textbackslash s}
sources it.
If there is e.g. code or notes in other projects I am working
on, I reach them through \tttt{<Alt-arrows>} because they
are in separate Vim instances
in Byobu/Tmux sessions and windows.
Because I stay for hours editing (\LaTeX and Python
scripts), I change the background to red, and eventually to green.
Because the blue color scheme does not highlight the spelling errors
if using true color, I run
\tttt{:hi SpellBad guifg=red guibg=lightblue}
to see the words found wrong by the spell checker.

\section{My vimrc file and usage}
In using Vim with my vimrc file~\cite{vimrc},
I mostly toggle the status line with \texttt{\textbackslash\textbackslash B}
and the tab line with \texttt{\textbackslash\textbackslash T}.
Save and close windows with \texttt{\textbackslash w} and \texttt{\textbackslash q}.
The mappings for transitioning through splits and tabs
are also used constantly.
Although the file is commented, one should
look for the help pages on the options that (s)he does
not understand, as a thorough explanation
of the file is tedious and out of the scope
of this document.

@@

\subsection{Normal mode}
Every letter and character in the keyboard is used.
In Normal mode: \ttt{TAB} is the same as \ttt{C-I},
\ttt{BS} and \ttt{C-H} are the same as h.
\ttt{C-J} and \ttt{C-N} are the same as j.
\ttt{C-P} is the same as k.
Space is the same as l.
\ttt{C-[} and \ttt{Esc} are not used,
\tttt{<C-\textbackslash> a-z} are reserved for extensions,
and \ttt{C-\_} is not used.
\tttt{+} is the same as \ttt{CR} and they are both not very useful.
Del is the same as x.

@@

Directions, home, end, page up and down, insert, all have mappings
in more centraly located keys.
The letters beggin unbounded mappings (require another character to
trigger an action without having to wait for \tttt{:se timeoutlen}): g,z,[].

@@

\subsection{Insert mode}
All standard character keys are used for entering text.
\tttt{<C-G>(j,k)} can be achieved by \tttt{<C-O>(j,k)} which is more powerful
in moving through multiple lines.
\ttt{C-$[$} is the same as \ttt{ESC}.
\ttt{C-J} and \ttt{C-M} are \ttt{CR}.
\tttt{<C-\textbackslash> a-z} are reserved for extensions,
other combinations with \ttt{C-\textbackslash} are not used.
